---
title: "Case Study: Benchmarking Algorithms using Eunomia"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmarking Algorithms using Eunomia}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  eval = FALSE
)
```

For this example, we are going to run a small experiment that compares the performance of the XGBoost algorithm to the performance of L1-penalised logistic regression. We are going to make use of the Eunomia dataset.

## Setup

We start by loading the packages that we will use. 

```{r setup}
library(Eunomia)
library(PLPBenchmarks)
library(xgboost)
```

Define the connectionDetails object for Eunomia:

```{r}
connectionDetails <- getEunomiaConnectionDetails()
```

Some other variables we need to define apriori:

```{r}
saveDirectory = "comparisonsVignette"
seed = 42
cdmDatabaseSchema = "main"
cdmDatabaseName = "Eunomia"
cdmDatabaseId = "Eunomia"
cohortDatabaseSchema = "main"
outcomeDatabaseSchema = "main"
cohortTable = "cohort"
```

We can have an overview of the pre-specified problems for Eunomia. We are going to compare the two algorithms on the following problem:

```{r}
data("eunomiaTasks")
eunomiaTasks$problemSpecification[1]
```

Let's load the benchmark design for the Eunomia prediction problems.

```{r}
data("eunomiaDesigns")
```

Let's continue by creating the cohorts we will work with.  

```{r}
Eunomia::createCohorts(connectionDetails = connectionDetails)
```

Define our database details. 

```{r}
databaseDetails <- PatientLevelPrediction::createDatabaseDetails(connectionDetails = connectionDetails, 
                                                                 cdmDatabaseSchema = cdmDatabaseSchema,
                                                                 cdmDatabaseName = cdmDatabaseName,
                                                                 cdmDatabaseId = cdmDatabaseId, 
                                                                 cohortDatabaseSchema = cohortDatabaseSchema,
                                                                 cohortTable = cohortTable,
                                                                 outcomeDatabaseSchema = outcomeDatabaseSchema,
                                                                 outcomeTable = cohortTable, 
                                                                 )
```

## Specifying our benchmark

We are going to set up our algorithm settings,

```{r}
lassoSettings <- PatientLevelPrediction::setLassoLogisticRegression(seed = seed)

xgbSettings <- PatientLevelPrediction::setGradientBoostingMachine(seed = seed)
```

and pass them to our model designs:

```{r}
selectedDesignList <- eunomiaDesigns[c(1, 1)]
names(selectedDesignList) <- c("GIBinCLXB_lasso", "GIBinCLXB_xgb")
names(selectedDesignList)
```

The `GIBinCLXB_lasso` model design already is using LASSO, to check:

```{r}
attr(selectedDesignList$GIBinCLXB_lasso$modelSettings$param, "settings")$name
```

We need to change the `modelSettings` for the second design that will execute the XGBoost algorithm. 

```{r}
selectedDesignList$GIBinCLXB_xgb$modelSettings <- xgbSettings
# Just to verify the algorithm has indeed changed:
attr(selectedDesignList$GIBinCLXB_xgb$modelSettings$param, "settings")$modelName
```

Let's create our benchmark design:

```{r}
comparisonBenchmark <- createBenchmarkDesign(modelDesign = selectedDesignList, 
                                                 databaseDetails = databaseDetails, 
                                                 saveDirectory = saveDirectory)
```

We can now have a view at our settings for the benchmark

```{r}
viewBenchmarkSettings(benchmarkDesign = comparisonBenchmark) %>%
  knitr::kable() %>%
  kableExtra::kable_paper(lightable_options = "striped") %>%
  kableExtra::scroll_box(width = "100%", height = "200px")
```

As we see, all settings between the two designs are equal, except the algorithm to be used, as we wanted.  
  
Now let's extract the data.  

```{r}
extractBenchmarkData(benchmarkDesign = comparisonBenchmark)
```

Finally, let's run our benchmark. 

```{r, results='hide'}
runBenchmarkDesign(benchmarkDesign = comparisonBenchmark)
```

And a let's have a look at some of the results.

```{r}
results <- PLPBenchmarks::getBenchmarkModelPerformance(benchmarkDesign = comparisonBenchmark)
results$performanceMetrics %>%
  dplyr::filter(metric %in% c("AUROC", "AUPRC", "calibrationInLarge mean prediction"))
```

