---
title: "Using PLPBenchmarks on Real-World Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PLPBenchmarks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  eval = FALSE
)
```

In this vignette, we will provide an introduction of how to make use of the `PLPBenchmarks` packages to build and execute benchmarks on real-world data (RWD). We will not show the whole pipeline as we need to be connected to a database. Working examples with the open-source `Eunomia` dataset exist in other vignettes (see vignette("BenchmarksWithEunomia") and vignette("AlgorithmComparisons"). 

## Step 1: Preparing the environment and specifying a benchmark design

To get started we load the `PLPBenchmarks` package and the other packages we will need. 

```{r setup}
library(DatabaseConnector)
library(PLPBenchmarks)
```

This loads the `PatientLevelPrediction` and `DatabaseConnector` R packages. 

We can see the problems for RWD, their specification and other necessary information to build the cohorts.

```{r, results='asis'}
data("tasks")
View(tasks)
```

We can also load the pre-defined model designs

```{r}
data("modelDesigns")
length(modelDesigns)
names(modelDesigns)
```

As we see, the pre-specified model designs is a list containing 13 `modelDesign` objects, one for each problem in the `tasks` data frame.

Before moving on, we need to define two more object.

The first object, is the `connectionDetails` from the `DatabaseConnector` package that we will use to connect to our database. For example, for a PostgreSQL database, we will define something like  

```{r}
conDets <- createConnectionDetails(dbms = "postgresql", 
                                   user = "user",
                                   password = "password",
                                   server = "server")
```

where `"user"`, `"password"` are the user credentials, and `"server"` the host name of the server and the relevant schemas. 

The second object we need to create is the `databaseDetails` from `PatientLevelPrediction`.

```{r}
dbDets <- createDatabaseDetails(connectionDetails = conDets, 
                                cdmDatabaseSchema = "cdmDatabaseSchema", 
                                cdmDatabaseName = "cdmDatabaseName", 
                                cdmDatabaseId = "cdmDatabaseId", 
                                cohortDatabaseSchema = "cohortDatabaseSchema", 
                                cohortTable = "cohortTable", 
                                outcomeDatabaseSchema = "outcomeDatabaseSchema", 
                                outcomeTable = "outcomeTable")
```

The names in quotes should be the corresponding schemas and table names that our target and outcome cohorts are located.

Now we can create a benchmark design.

```{r}
benchmarkDesign <- createBenchmarkDesign(modelDesign = modelDesigns, databaseDetails = dbDets, saveDirectory = "gettingStarted")
```

Note the `saveDirectory` argument takes a character or a file path for the location to save the results. 

> **Some more details:**   
The function `createBecnhmarkDesign()` takes three arguments: a list of named model designs, a `databaseDetails` object created from  `PatientLevelPrediction::createDatabaseDetails()` function, and a name for the directory to save the results. The returned object is still a list, but this time a list of class `benchmarkDesign` with the additional information provided in each list element (i.e. each model design for each problem). The nice thing about this function, is identifying the unique sets of covariates and study population settings that are needed to be created in a benchmark design. These information is stored in the attributes of the returned object.   
The advantage of this is that for the model designs that share common target and outcome cohorts, along with common covariates and population settings, the `plpData` and `studyPopulation` objects will only be created once, saving time from recreating them each time we are running a model design. 


We can now inspect our benchmark design settings in a side-to-side comparison to check if everything is specified correctly.

```{r, results='asis'}
viewBenchmarkSettings(benchmarkDesign = benchmarkDesign) %>%
  knitr::kable() %>%
  kableExtra::kable_paper(lightable_options = "striped") %>%
  kableExtra::scroll_box(width = "100%", height = "200px")
```

If everything is specified correctly and we are ready to go we can then proceed. 

## Step 2: Create Cohorts and extract covariates (and even create the study population)

Next, we create the cohorts in the database

```{r}
createBenchmarkCohorts(benchmarkDesign = benchmarkDesign)
```

We are now ready to proceed.   
  
Last think before running the benchmark design, is to extract the data (covariates) that are defined in each model design.  

```{r}
extractBenchmarkData(benchmarkDesign = benchmarkDesign, createStudyPopulation = TRUE)
```

Notice that the function is responsible for two things:
 - extracts the covariates that are needed for each model designs
 - optionally creates the study population (default to TRUE)
This is by convention to provide flexibility on altering the states of covariates and population if needed. 

## Step 3: Run the benchmark design

```{r}
runBenchmarkDesign(benchmarkDesign = benchmarkDesign)
```

`runBencmarkDesign()` which will run all specified model designs. 

## Step 4: Inspect and share results

The following will collect all results in one object called `results`.

```{r}
results <- getBenchmarkModelPerformance(benchmarkDesign = benchmarkDesign)
```

The `results` object is a list with two components: a `performanceMetrics` data frame and an `executionTimes` data frame. 
Optionally, we can also call 

```{r}
viewBenchmarkResults(benchmarkDesign = benchmarkDesign, viewShiny = TRUE)
```

to create and view a shiny app with the results. 

For more complete examples, have a look at the other vignettes using the Eunomia dataset to simulate experiments. 





