---
title: "Using PLPBenchmarks on Real-World Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PLPBenchmarks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette, we will provide an introduction of how to make use of the `PLPBenchmarks` packages to build and execute benchmarks on real-world data (RWD). We will not show the whole pipeline as we need to be connected to a database. Working examples with the open-source `Eunomia` dataset exist in other vignettes (see vignette("BenchmarksWithEunomia") and vignette("AlgorithmComparisons"). 

To get started we load the `PLPBenchmarks` package and the other packages we will need. 

```{r setup}
library(DatabaseConnector)
library(PLPBenchmarks)
```

This loads the `PatientLevelPrediction` and `DatabaseConnector` R packages. 

We can see the problems for RWD, their specification and other necessary information to build the cohorts.

```{r, results='asis'}
data("tasks")
tasks %>%
  knitr::kable() %>%
  kableExtra::kable_classic(lightable_options = "striped") %>%
  kableExtra::scroll_box(width = "100%", height = "200px")
```

We can also load the pre-defined model designs

```{r}
data("modelDesigns")
length(modelDesigns)
names(modelDesigns)
```

As we see, the pre-specified model designs is a list containing 13 `modelDesign` objects, one for each problem in the `tasks` data frame.

Before moving on, we need to define two more object.

The first object, is the `connectionDetails` from the `DatabaseConnector` package that we will use to connect to our database. For example, for a PostgreSQL database, we will define something like  

```{r}
conDets <- createConnectionDetails(dbms = "postgresql", 
                               user = "user",
                               password = "password",
                               server = "server")
```

where `"user"`, `"password"` are the user credentials, and `"server"` the host name of the server and the relevant schemas. 

The second object we need to create is the `databaseDetails` from `PatientLevelPrediction`.

```{r}
dbDets <- createDatabaseDetails(connectionDetails = conDets, 
                                cdmDatabaseSchema = "cdmDatabaseSchema", 
                                cdmDatabaseName = "cdmDatabaseName", 
                                cdmDatabaseId = "cdmDatabaseId", 
                                cohortDatabaseSchema = "cohortDatabaseSchema", 
                                cohortTable = "cohortTable", 
                                outcomeDatabaseSchema = "outcomeDatabaseSchema", 
                                outcomeTable = "outcomeTable")
```

The names in quotes should be the corresponding schemas and table names that our target and outcome cohorts are located.

Now we can create a benchmark design.

```{r}
benchmarkDesign <- createBenchmarkDesign(modelDesign = modelDesigns, databaseDetails = dbDets, saveDirectory = "gettingStarted")
```

Note the `saveDirectory` argument takes a character or a file path for the location to save the results. 

We can now inspect our benchmark design settings in a side-to-side comparison to check if everything is specified correctly.

```{r, results='asis'}
viewBenchmarkSettings(benchmarkDesign = benchmarkDesign) %>%
  knitr::kable() %>%
  kableExtra::kable_paper(lightable_options = "striped") %>%
  kableExtra::scroll_box(width = "100%", height = "200px")
```

If everything is specified correctly and we are ready to go we can then call `runBencmarkDesign()` which will run all specifed model designs. For more complete example, have a look at the other vignettes. 





